---
title: "Automatic Thoughts Questionnaire"
author: "MochiBear.Hei"
date: "09.05.2024"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```

```{r setup,}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

# Data manipulation
library(readxl)
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(reshape2)

# Plotting
library(ggplot2)
library(naniar)

# Factor analysis
library(psych)
library(lavaan)

# Random Forest visualizations
library(randomForest)
library(caret)
```

# This script loads and preprocesses ATQ assessment data and consent records.
Key steps include:
- Renaming and selecting relevant variables
- Standardizing assessment context labels and treatment names
- Merging with consent data and filtering by treatment type
- Handling missing data and duplicates
- Creating missing "Admission" rows using "Assessment" data
- Standardizing the ATQ sum score and renaming timepoints
- Final dataset is saved as 'atq_f.csv' for downstream modeling

# Notable: Removes respondents with >30% missing data, ensures all have post-treatment data,
and imputes missing Admission scores based on Assessment values.

# Read in data
```{r, warning=FALSE, message=FALSE}
ATQ <- read_excel("/Users/maggieheimvik/Desktop/COPE/data/dataset/ATQ_avid.xls") #3710 obs. of 57 var
#View(ATQ)
  
consent <- read_csv("/Users/maggieheimvik/Desktop/COPE/data/dataset/scripts/anon/consent_a.csv") #5901 obs. of 30 var
#View(consent)

```

### Change names

```{r}
# Check names
colnames(ATQ)

# Rename
ATQ <- ATQ %>% 
  rename(
    assessment_id = `assessment instance id`,
    assessment_version = `assessment version`,
    assessment_title = `assessment instance title`,
    assessment_owner_id = `assessment instance owner id`,
    assessment_start_date = `assessment instance start date`,
    assessment_end_date = `assessment instance end date`,
    assessment_created_date = `assessment instance created date`,
    assessment_last_modified = `assessment instance last modified/submitted`,
    assessment_started = `assessment instance has started`,
    assessment_submitted = `assessment instance is submitted`,
    assessment_closed = `assessment instance is closed`,
    assessment_context_label = `assessment instance context label`,
    assessment_first_started_date = `assessment instance first time started date`,
    assessment_first_submitted_date = `assessment instance first time submitted date`,
    assessment_portal = `assessment instance portal of submission`,
    treatment_id = `treatment id`,
    treatment_name = `treatment name`,
    treatment_type_id = `treatment type id`,
    treatment_type_name = `treatment type name`,
    treatment_role = `treatment role`,
    respondent_id = `respondent id`,
    respondent_username = `respondent username`,
    respondent_account_enabled = `respondent account enabled`,
    respondent_test_account = `respondent test account`,
    respondent_last_login = `respondent last login`,
    respondent_communcation_disabled = `respondent communication disabled`, 
    Q1 = `Q1`,
    Q2 = `Q2`,
    Q3 = `Q3`,
    Q4 = `Q4`,
    Q5 = `Q5`,
    Q6 = `Q6`,
    Q7 = `Q7`,
    Q8 = `Q8`,
    Q9 = `Q9`,
    Q10 = `Q10`,
    Q11 = `Q11`,
    Q12 = `Q12`,
    Q13 = `Q13`,
    Q14 = `Q14`,
    Q15 = `Q15`,
    Q16 = `Q16`,
    Q17 = `Q17`,
    Q18 = `Q18`,
    Q19 = `Q19`,
    Q20 = `Q20`,
    Q21 = `Q21`,
    Q22 = `Q22`,
    Q23 = `Q23`,
    calc_aa = `calculation:MODUMBAD-ATQ-AA`,               
    calc_pan = `calculation:MODUMBAD-ATQ-PAN-AA`,            
    calc_pan_sum = `calculation:MODUMBAD-ATQ-PAN-SUM`,           
    calc_ptsd = `calculation:MODUMBAD-ATQ-PTSD-AA`,           
    calc_ptsd_sum = `calculation:MODUMBAD-ATQ-PTSD-SUM`,        
    calc_sos = `calculation:MODUMBAD-ATQ-SOS-AA`,         
    calc_sos_sum = `calculation:MODUMBAD-ATQ-SOS-SUM`,          
    calc_sum = `calculation:MODUMBAD-ATQ-SUM`
 )

```

Selecting only necessary columns
```{r}
# Select only from the BAI dataset
ATQ <- ATQ %>% 
  select(
    respondent_id, 
    assessment_context_label,
    assessment_created_date,
    treatment_id,
    treatment_name, 
    treatment_type_id, 
    assessment_created_date,
    starts_with("Q"), 
    starts_with("calc")
  )

# Select only in the consent
consent<- consent %>% 
  select(respondent_id, consent)
```
